Quantum Mechanics Assignment Five

### 2016-02-12 quantum/latex/a5.tex
bla

\textbf{b.)} Now we apply a 0.5 V/m electric field and see how the spectrum changes. The Stark
Effect tells us that the energy levels will be perturbed by either zero or
$\pm3e\epsilon a_0$, which for this electric field would be $\pm 2.64*10^{-2}eV$. This comes out
to be a change of roughly $0.25$ percent, which is a fairly small value.\\

@@

@@

\section{11.9: Commutators}
Here we show that the spin magnitude operators \textbf{$S^2$} and \textbf{$I^2$} commute with
the perturbation Hamiltonian, while the component operators \textbf{$S_z$} and \textbf{$I_z$}
do not.

\begin{align*}
  \left[S^2, H'_{hf}\right] &= \left(S^2H'_{hf} - H'_{hf}S^2\right)\ket{m_p,m_s}\\
  &= S^2H'_{hf}\ket{m_p,m_s} - H'_{hf}S^2\ket{m_p,m_s}\\
  &= S^2E_{m_p,m_s}\ket{m_p,m_s} - H'_{hf}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
  &= \frac{3\hbar^2}{4}E_{m_p,m_s}\ket{m_p,m_s} - E_{m_p,m_s}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
  &= 0\\
\end{align*}

>>>>

\textbf{b.)} Now we apply a 0.5 V/m electric field and see how the spectrum changes. The Stark
Effect tells us that the energy levels will be perturbed by either zero or
$\pm3e\epsilon a_0$, which for this electric field would be $\pm 7.94*10^{-2}eV$. This comes out
to be a change of roughly $0.77$ percent, which is a fairly small value.\\

@@

\section{11.2: Proving raising / lowering operators}
Here we show that the raising and lowering operators for angular momentum behave similarly
to the QHO raising and lowering operators.\\

First, we must show that the operators raise and lower the z-component momentum measurement
by one unit:\\

\begin{align*}
  J_z(J_+\ket{j,m_j}) &= (\hbar J_+ + J_+J_z)\ket{j,m_j}\\
  &= \hbar J_+\ket{j,m_j} + m_j \hbar J_+\ket{j,m_j}\\
  J_z(J_+\ket{j,m_j}) &= (m_j+1)\hbar(J_z\ket{j,m_j})
  \hspace{2cm}\mbox{J+ raises energy by $\hbar$}\\
\end{align*}

\begin{align*}
  J_z(J_-\ket{j,m_j}) &= (-\hbar J_- + J_-J_z)\ket{j,m_j}\\
  &= \hbar J_-\ket{j,m_j} + m_j \hbar J_-\ket{j,m_j}\\
  J_z(J_-\ket{j,m_j}) &= (m_j-1)\hbar(J_z\ket{j,m_j})
  \hspace{2cm}\mbox{J- lowers energy by $\hbar$}\\
\end{align*}

This shows that the operation of $J_\pm$ on a state before the operation of $J_z$ effectively
raises or lowers the measurement of that state by $\hbar$. We now need to find the relationship
between $J_\pm\ket{j,m_j}$ and $\ket{j,m_j\pm1}$.\\

\begin{align*}
  |J_+\ket{j,m_j}|^2 &= c^2\\
  &= \bra{j,m_j}J_-J_+\ket{j,m_j}\\
  &= \bra{j,m_j}(J_x-iJ_y)(J_x+iJ_y)\ket{j,m_j}\\
  &= \bra{j,m_j}J_xJ_x + J_yJ_y + i[J_x,J_y]\ket{j,m_j}\\
  &= \bra{j,m_j}J^2 - J_zJ_z + i[J_x,J_y]\ket{j,m_j}\\
  &= \bra{j,m_j}J^2 - J_zJ_z - \hbar J_z\ket{j,m_j}\\
  &= j(j+1)\hbar - m_j^2\hbar^2 - m_j\hbar^2\\
  &= \hbar^2\left(j(j+1)\hbar - m_j(m_j + 1)\right)\\
  &\rightarrow J_+\ket{j,m_j} = \hbar\sqrt{j(j+1) - m_j(m_j+1)}\ket{j,m_j+1}\\
\end{align*}

\begin{align*}
  |J_-\ket{j,m_j}|^2 &= c^2\\
  &= \bra{j,m_j}J_+J_-\ket{j,m_j}\\
  &= \bra{j,m_j}(J_x+iJ_y)(J_x-iJ_y)\ket{j,m_j}\\
  &= \bra{j,m_j}J_xJ_x + J_yJ_y - i[J_x,J_y]\ket{j,m_j}\\
  &= \bra{j,m_j}J^2 - J_zJ_z - i[J_x,J_y]\ket{j,m_j}\\
  &= \bra{j,m_j}J^2 - J_zJ_z + \hbar J_z\ket{j,m_j}\\
  &= j(j+1)\hbar^2 - m_j^2\hbar^2 + m_j\hbar&2\\
  &= \hbar^2\left(j(j+1) - m_j(m_j-1)\right)\\
  &\rightarrow J_-\ket{j,m_j} = \hbar\sqrt{j(j+1) - m_j(m_j-1)}\ket{j,m_j-1}\\
\end{align*}

We've just shown that the angular momentum ladder operators behave the same way the quantum
harmonic oscillator operators do. This is powerful, since it will allow us to simplify math
later on. It is important to note that the raising and lowering operators only act on the
Z-component of spin, and that they have two termination conditions: at the minimum and
maximum Z-component.\\

\textit{Inspired by Jared Cayton}\\

@@

\section{11.8: Hyperfine Hamiltonian}
Here we use ladder operators to find the hyperfine Hamiltonian for the Hydrogen atom:\\

\begin{align*}
  H_{hf}' &=
  \begin{pmatrix}
    \bra{++}H_{hf}'\ket{++} &
    \bra{++}H_{hf}'\ket{+-} &
    \bra{++}H_{hf}'\ket{-+} &
    \bra{++}H_{hf}'\ket{--} &\\
    \bra{+-}H_{hf}'\ket{++} &
    \bra{+-}H_{hf}'\ket{+-} &
    \bra{+-}H_{hf}'\ket{-+} &
    \bra{+-}H_{hf}'\ket{--} &\\
    \bra{-+}H_{hf}'\ket{++} &
    \bra{-+}H_{hf}'\ket{+-} &
    \bra{-+}H_{hf}'\ket{-+} &
    \bra{-+}H_{hf}'\ket{--} &\\
    \bra{--}H_{hf}'\ket{++} &
    \bra{--}H_{hf}'\ket{+-} &
    \bra{--}H_{hf}'\ket{-+} &
    \bra{--}H_{hf}'\ket{--} &\\
  \end{pmatrix}\\
  &=
  \frac{A}{\hbar^2}
  \begin{pmatrix}
    \bra{++}\vec{S}\cdot\vec{I}\ket{++} &
    \bra{++}\vec{S}\cdot\vec{I}\ket{+-} &
    \bra{++}\vec{S}\cdot\vec{I}\ket{-+} &
    \bra{++}\vec{S}\cdot\vec{I}\ket{--} &\\
    \bra{+-}\vec{S}\cdot\vec{I}\ket{++} &
    \bra{+-}\vec{S}\cdot\vec{I}\ket{+-} &
    \bra{+-}\vec{S}\cdot\vec{I}\ket{-+} &
    \bra{+-}\vec{S}\cdot\vec{I}\ket{--} &\\
    \bra{-+}\vec{S}\cdot\vec{I}\ket{++} &
    \bra{-+}\vec{S}\cdot\vec{I}\ket{+-} &
    \bra{-+}\vec{S}\cdot\vec{I}\ket{-+} &
    \bra{-+}\vec{S}\cdot\vec{I}\ket{--} &\\
    \bra{--}\vec{S}\cdot\vec{I}\ket{++} &
    \bra{--}\vec{S}\cdot\vec{I}\ket{+-} &
    \bra{--}\vec{S}\cdot\vec{I}\ket{-+} &
    \bra{--}\vec{S}\cdot\vec{I}\ket{--} &\\
  \end{pmatrix}\\
   = & \frac{A}{\hbar^2}
  \begin{pmatrix}
    \frac{\hbar^2}{4}&
    \bra{++}\frac{1}{2}S_-I_+\ket{+-}&
    \bra{++}\frac{1}{2}S_+I_-\ket{-+}&
    0&\\
    0&
    \bra{+-}\frac{1}{2}S_-I_+\ket{+-} - \frac{\hbar^2}{4}&
    \bra{+-}\frac{1}{2}S_+I_-\ket{-+}&
    0&\\
    0&
    \bra{-+}\frac{1}{2}S_-I_+\ket{+-}&
    \bra{-+}\frac{1}{2}S_+I_-\ket{-+} - \frac{\hbar^2}{4}&
    0&\\
    0&
    \bra{--}\frac{1}{2}S_-I_+\ket{+-}&
    \bra{--}\frac{1}{2}S_+I_-\ket{-+}&
    \frac{\hbar^2}{4} &\\
  \end{pmatrix}\\
  = & \frac{A}{\hbar^2}
  \begin{pmatrix}
    \frac{\hbar^2}{4}&
    0&
    0&
    0&\\
    0&
    -\frac{\hbar^2}{4}&
    \frac{2\hbar^2}{4}&
    0&\\
    0&
    \frac{2\hbar^2}{4}&
    -\frac{\hbar^2}{4}&
    0&\\
    0&
    0&
    0&
    \frac{\hbar^2}{4} &\\
  \end{pmatrix}\\
  = & \frac{A}{4}
  \begin{pmatrix}
    1&
    0&
    0&
    0&\\
    0&
    -1&
    2&
    0&\\
    0&
    2&
    -1&
    0&\\
    0&
    0&
    0&
    1&\\
  \end{pmatrix}\\
\end{align*}

We have succesfully shown how expanding the angular momentum dot product and using ladder
operators allows each element in our perturbation Hamiltonian to be solved for. Our end
product is a nondiagonal matrix with first and second-order corrections. Unfortunately our
original energies are degenerate, as this is the n=2 state. Thus we must diagonalize this
matrix to get a new diagonal perturbation matrix representing the changes to our original
system.\\
